Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-06 paper/paper.tex
fix: PC hypothesis 1.0

\subsection{Predictive coding hypothesis}
The predictive coding framework
\citep{clark2013whatever, friston2008hierarchical}
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

\begin{itemize}
  \item Both the predictive coding account and POMDPs
  are process theories backed up by plausible
  neurobiological evidence that
  view the brain as a ``statistical organ''
  generalizing from the past to new experiences.
  \item Both provide a parsimonious explanation why the
  human brain decreases processing load devoted to incoming information
  when the environment is predictable and allocate increased
  attentional resources when novel stimuli are encountered.
  Similarly, both propose explanations why
  predicting the future is inherently
  linked to information from the past.
  \item Sensory experience is a generative process from both views.
  In predictive coding, sensory perception of the external world
  is a generative experience due to modulatory top down experience at
  various hierarchical levels of sensory input processing.
  In our RL view, the (partially observed)
  environmental model is incorporated in the POMDP,
  \textit{which can be fully recovered based on the last
  state alone}.
  \item Both naturally expose a mechanisms of brain plasticity in that
  neuronal wiring gets increasingly adapted
  when faced by unprecedented environmental
  challenges.
  \item The hierarchical aspects from predictive coding
  is re-expressed in POMDPs in form of
  nested prediction of probable upcoming actions and rewards.
  \item Both model the consequences of action. In RL, the horizon of that
  look into the future is explicitly manifested in the $\gamma$ parameter
  in the Bellman equation, but not
  explicitly modeled in the PC account that
  tends to emphasize prediction about the
  immediate future.
  \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
  by the difference between actually predicted reward of an action given
  a state and the reward predicted by non-linear regression with graded
  discounting of future rewards.}
  \item Adapting the neuronal connections for improved top-down modulation
  takes the concrete form of gradient computation
  \textit{and back-propagation} in MDPs and RL,
  although the neurobiological plausibility of
  the back-propagation procedure is controversal
  \citep{goodfellow2016deep}.
  \item In sum,
  the MDP account may serve as a concrete conceptualization of the PC account
  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
  mechanism of the horizon of future considerations or
  how the internal model of the world gets updated,
  and collapses sensory input processing and action output preparation.
  %
  As such, our MDP account of DMN function can parsimoneously explain
  prediction coding hypothesis.
\end{itemize}

>>>>

\subsection{Predictive coding hypothesis}
The predictive coding account
\citep{clark2013whatever, friston2008hierarchical}
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

There a number of common properties between the predictive coding account
and the proposed formal account.
Both frameworks are plausibilized by neurobiological evidence that
the human brain may be viewed as a ``statistical organ'' aiming to
generalize from the past to new experiences.
Both accounts provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable and allocate increased
computation resources when novel stimuli or unexpected events are encountered.
The predictive coding and formal account
naturally evoke a mechanisms of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental
challenges.

While sensory experience is a generative process from both views,
in predictive coding sensory perception of the external world
is a generative experience due to modulatory top down-experience at
various hierarchical levels of sensory input processing. Instead,
the Markov property of our sequential decision making framework
In our RL view, the environmental model is
\textbf{effectively expressed in the current state of the human agent
XXX guys, please double-check}.
The hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both model the consequences of action.
In the present account, the horizon of that
look into the future is explicitly expressed in the $\gamma$ parameter
in the Bellman equation.
The predictive coding framework is typically explained without
explicitly parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
The process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation
back-propagation in our RL implementation,
although the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.
\item In sum,
the formal account may serve as a concrete implementation of
the predictive coding account
from cognitive neuroscience.
MPDs have the advantage of exposing an explicit
mechanism of the horizon of future considerations or
how the internal model of the world gets updated,
and collapses sensory input processing and action output preparation.
%
As such, recasting DMN function in terms of MDPs naturally incorporates
a majority of aspects of the prediction coding hypothesis.
